### 2024-03-10 manuscript.Rmd
Start working on the project

---
title             : "Formal Model of the Set-Point Theory of Subjective Well-Being"
shorttitle        : "Set-Point Theory of Subjective Well-Being"

author: 
  - name          : "Alp Kaan Aksu"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Leopoldstraße 13"
    email         : "a.aksu@campus.lmu.de"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"

affiliation:
  - id            : "1"
    institution   : "LMU Munich"

authornote: |
  Matriculation Number: 12607598

abstract: |
  There is research that shows...

  Here, we model the set point theory of subjective well being

keywords          : ["happiness", "subjective well-being", "hedonic treadmill", "set-point theory"]
wordcount         : "X"

bibliography      : "r-references.bib"

floatsintext      : no
linenumbers       : no
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")

source("VAST/VAST.R")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r construct-definitions, include=FALSE}
swb <- construct("SWB")
swb_set_point <- construct("SWBSetPoint")
adaptation_rate <- construct("AdaptationRate")
event <- construct("Event")
```

# Deneme
```{r vast-display}
VAST(swb, swb_set_point, adaptation_rate, event) |>
  rel(swb_set_point %-c->% swb) |>
  rel(adaptation_rate %-c->% swb) |>
  rel(event %-c->% swb) |>
  analysts("Alp") |> fimm() |>
  render_vast()
```

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.

# Results

# Discussion

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::

### 2024-03-10 VAST/VAST.R

@@ -1,0 +1,157 @@
+library(DiagrammeR)
+
+construct <- function(id) {
+  list(
+    id = id,
+    names = c()
+  )
+}
+
+VAST <- function(...) {
+  construct_params <- list(...)
+  list(
+    constructs = construct_params,
+    relationships = list(),
+    analysts = list(),
+    fimm = FALSE
+  )
+}
+
+find_node_by_type <- function (graph, type_id) {
+  get_node_ids(graph, conditions = type == type_id)
+}
+
+render_namings <- function (vast) {
+  diagram <- create_graph()
+  
+  for (construct in vast$constructs) {
+    diagram <- diagram |>
+      add_node(
+        label = construct$id,
+        type = construct$id,
+        node_aes = node_aes(
+          fixedsize = FALSE, 
+          shape = "rectangle",
+          fillcolor = "#ffffff"
+        )
+      )
+    
+    for (name in construct$names) {
+      diagram <- diagram |>
+        add_node(
+          label = paste0("'", name, "'"),
+          type = paste0(construct$id, "_name"),
+          node_aes = node_aes(
+            fixedsize = FALSE, 
+            shape = "rectangle",
+            fillcolor = "#ffffff"
+          )
+        )
+      
+      ids <- get_node_ids(diagram)
+      id <- ids[length(ids)]
+      
+      print(id)
+      
+      diagram <- diagram |>
+        add_edge(
+          from = find_node_by_type(diagram, construct$id),
+          to = id,
+          edge_aes = edge_aes(
+            label = "n"
+          )
+        )
+    }
+    
+  }
+  
+  render_graph(diagram, layout = "nicely")
+}
+
+render_vast <- function (vast) {
+  diagram <- create_graph()
+  
+  for (construct in vast$constructs) {
+    diagram <- diagram |>
+      add_node(
+        label = construct$id,
+        type = construct$id,
+        node_aes = node_aes(
+          fixedsize = FALSE, 
+          shape = "rectangle",
+          fillcolor = "#ffffff"
+        )
+      )
+  }
+  
+  for (relationship in vast$relationships) {
+    diagram <- diagram |>
+      add_edge(
+        from = find_node_by_type(diagram, relationship$from),
+        to = find_node_by_type(diagram, relationship$to),
+        edge_aes = edge_aes(
[…]

### 2024-03-16 manuscript.Rmd
Add introduction and robustness sections

@@
 
-bibliography      : "r-references.bib"
+bibliography      : ["r-references.bib", "references.bib"]
 
-floatsintext      : no
+floatsintext      : yes
 linenumbers       : no
@@
 
-```{r construct-definitions, include=FALSE}
-swb <- construct("SWB")
-swb_set_point <- construct("SWBSetPoint")
-adaptation_rate <- construct("AdaptationRate")
-event <- construct("Event")
+# Introduction
+
+One of the most important questions we might attempt to answer is how to live our lives and what makes life good.
+Trying to maximize our happiness is the naive answer we easily arrive at, we want to feel good, but *happiness* is an ambiguous word.
+Although it seems like all people tend to evaluate their lives and think about *happiness* [@veenhoven11HowUniversal2010], the understanding of *happiness*, its use and connotations vary across time and place [@oishiConceptsHappinessTime2013].
+@dienerSubjectiveWellbeing1984 suggests the use of a scientific term: *subjective well-being* (SWB).
+
+## Subjective Well-Being
+According to @dienerSubjectiveWellbeing1984, SWB is the subjective global evaluation of life and consists of three components: (1) life satisfaction, (2) positive affect, (3) negative affect.
+
+## The Phenomenon: Hedonic Treadmill
+
+The observation, that people go back to a stable level of SWB after (major) life events is not new.
+@helsonAdaptationlevelTheoryExperimental1964 formulated the more general Adaptation Level Theory, stating that how people perceive stimuli is influenced by their past experiences.
+@brickman1971adaptation applied this idea to SWB for the first time and coined the term *hedonic treadmill*, which is still commonly used to refer to the phenomenon of returning to a base SWB level after life events [e.g.
+@czaikaMigrationAspirationsAre2014, manciniSteppingHedonicTreadmill2011], and @brickmanLotteryWinnersAccident provided first evidence for it.
+
+## Our Goal
+
+We aim to formalize the Set-Point Theory of SWB as described by @dienerHedonicTreadmillRevising2006a, which explains the phenomenon of hedonic treadmill and tries to account for individual differences.
+
+# Robustness of the phenomenon
+
+@luhmannSubjectiveWellbeingAdaptation2012 has done a meta-analysis on adaptation to life events focusing on SWB.
+Evaluating the robustness of the studies focusing on the hedonic treadmill is difficult, since no randomized controlled trials are possible.
+
+## Empirical Evidence
+
+According to @luhmannSubjectiveWellbeingAdaptation2012, there is unequivocal evidence that the hedonic treadmill is a robust phenomenon.
+
+## Generalizability
+
+We use the UTOS-Framework by @cronbach1982designing to evaluate the generalizability of the findings.
+
+### Units
+Qualitatively different samples.
+
+### Treatment
+No randomized controlled trial. But focus on lots of different major life events.
+
+### Outcome
+Using affective (AWB) and cognitive well being (CWB) measure (for some life events), studies used different measures for different variable, e.g.
+for the measurement of CWB in the context of marriage, relationship satisfaction was used.
+
+### Setting
+A lot of studies were conducted by the same team. Most of the data comes from the German Socioeconomical Panel.
+
+All in all, there are reasonable evidence for generalizability of the Hedonic Treadmill phenomenon.
+
+# The Prototheory
+
+| Aspect | Generali
+
+# Model
+
+The central ideao of
+
+# Implementation
+
+We made small changes to the theoretical model, in order to make it suitable for an implementation. We are making the following assumptions:
+
+- The construct of *Life Event* can be reduced to *Appraisal of the Life Event*, since only the latter is a part of the (following) relevant psychological processes.
+
+*VAST display of the model here*
+
+```{r child="./material/variable_definitions.Rmd"}
 ```
-
-# Deneme
-```{r vast-display}
-VAST(swb, swb_set_point, adaptation_rate, event) |>
-  rel(swb_set_point %-c->% swb) |>
-  rel(adaptation_rate %-c->% swb) |>
-  rel(event %-c->% swb) |>
-  analysts("Alp") |> fimm() |>
-  render_vast()
-```
-
-# Methods
-We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.
-<!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->
-
-## Participants
-
-## Material
-
-## Procedure
-
-## Data analysis
-We used `r cite_r("r-references.bib")` for all our analyses.
-
-# Results
 

### 2024-03-16 material/variable_definitions.Rmd
Add introduction and robustness sections

| **Construct** | **Scale Level** | **Range** | **Anchors** |
| --------- | ----------- | ----- | ------- |
| Subjective Well-Being | continuous | [0; 1] | 0 = absolutely not satisfied with life \newline 1 = maximal life satisfaction |
| Subjective Well-Being Set Point | continuous | [0; 1] | 0 = absolutely not satisfied with life \newline 1 = maximal life satisfaction |
| Appraisal of Life Event | continuous | [-1; 1] | -1 = negative \newline 0 = neutral \newline 1 = positive |

### 2024-03-17 manuscript.Rmd
Continue working on the theory

@@
 @luhmannSubjectiveWellbeingAdaptation2012 has done a meta-analysis on adaptation to life events focusing on SWB.
-Evaluating the robustness of the studies focusing on the hedonic treadmill is difficult, since no randomized controlled trials are possible.
+Evaluating the robustness of the studies focusing on SWB and adaptation is difficult, since it is impossible to manipulate life events experimentally and cross-sectional studies have a lack of internal validity [@luhmannSubjectiveWellbeingAdaptation2012].
+All of the following statements are based on the meta-analysis by @luhmannSubjectiveWellbeingAdaptation2012.
 
@@
 
-According to @luhmannSubjectiveWellbeingAdaptation2012, there is unequivocal evidence that the hedonic treadmill is a robust phenomenon.
+Studies focusing on the hedonic treadmill can not be adequatly characterized by single effect sizes, since the temporal development of SWB is of interest.
+So @luhmannSubjectiveWellbeingAdaptation2012 used a method that takes time into account.
+For each life event, they fitted a logarithmic model using the effects sizes reported in original studies.
+An adaptation is characterized by a change parameter, which is in the opposite direction of the initial reaction.
+For example, having a positive reaction (higher SWB) after a life event, then a negative change in SWB, indicates that an adaptation is taking place.
+
+Adaptations were observed for all life events, which means that the hedonic treadmill can be considered a robust phenomenon, but reduced internal validity due to the design of these studies makes it hard to evaluate.
 
@@
 
-We use the UTOS-Framework by @cronbach1982designing to evaluate the generalizability of the findings.
+We evaluate the generalizability of the findings guided by the UTOS framework  [@cronbach1982designing].
+A strict application of the UTOS framework to the phenomenon at hand is not possible because of the non-experimental nature of the studies.
 
 ### Units
-Qualitatively different samples.
+In a lot of studies, the majority of the participants were females (65,4% across all events; for some life events like bereavement and divorce even 70+).
+More than half of the studies were from the US.
 
 ### Treatment
-No randomized controlled trial. But focus on lots of different major life events.
+Studies focus on different major life events, 8 of them were included in the meta-analysis.
 
 ### Outcome
-Using affective (AWB) and cognitive well being (CWB) measure (for some life events), studies used different measures for different variable, e.g.
+Using affective (AWB) and cognitive well being (CWB) measure (for some life events), studies used different measures for different variables, e.g.
 for the measurement of CWB in the context of marriage, relationship satisfaction was used.
@@
 ### Setting
-A lot of studies were conducted by the same team. Most of the data comes from the German Socioeconomical Panel.
+aedasd
 
-All in all, there are reasonable evidence for generalizability of the Hedonic Treadmill phenomenon.
+Participants from diverse backgrounds are necessary for generalizability.
+Mixed (positive or negative) set of life event were analyzed, which is a good first sign for generalizability to different life events.
 
@@
 
-| Aspect | Generali
+## Constructs
+
+*TODO*
+
+- SWB: Definiton by Diener (include subtypes in the theoretical model but not in)
+  - AWB:
+  - CWB:
+- SWBSetPoint: SWB people return to
+- AdaptationRate: How fast the adaptation happens
+  - AdaptationRatePerson: Personenvariable
+  - AdaptationRateEvent: Hängt von InitialReaction ab
+- InitialReaction: Change in SWB directly after / at the time of the LifeEvent
+- LifeEvent: Status change
 
@@
 
-The central ideao of
+The central idea of
 
@@
 
-- The construct of *Life Event* can be reduced to *Appraisal of the Life Event*, since only the latter is a part of the (following) relevant psychological processes.
+- The constructs of *Life Event* and *Initial Reaction* can be reduced to *Initial Reaction*, since only the latter is a part of the (following) relevant psychological processes.
+-
 

### 2024-03-17 material/statements.Rmd
Continue working on the theory

| Identifier | Citation                                                                                                                                                                                                                                                                                                                                                                                                                                                                                               | Reference                                                            | Relationship?                      | Comments?                                                                      |     |
| ---------- | ------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------ | -------------------------------------------------------------------- | ---------------------------------- | ------------------------------------------------------------------------------ | --- |
| A          | A notable finding was that the initial reaction and the rate of adaptation varied considerably between different life events.                                                                                                                                                                                                                                                                                                                                                                          | Luhmann et al. (2012)                                                | Appraisal  - c → Adaptation Rate   | Valence of the life event changes the adaptation rate.                         |     |
| A1         | For instance, the initial reaction to marriage was positive, but subsequent adaptation was fast and completed after 2 years on average (Lucas et al., 2003).                                                                                                                                                                                                                                                                                                                                           |                                                                      |                                    | positive valence: fast adaptation                                              |     |
| A2         | In contrast, the rate of adaptation was much slower for negative events such as the onset of disability (Lucas, 2007b), widowhood (Lucas et al., 2003; Specht, Egloff, & Schmukle, 2011), and divorce (Lucas, 2005).                                                                                                                                                                                                                                                                                   |                                                                      |                                    | negative valence: slow adaptation                                              |     |
| A3         | The findings were perhaps most striking for unemployment: After the beginning of unemployment, mean levels of SWB were significantly below baseline, even if people became reemployed (Lucas et al., 2004). Repeated unemployment spells aggravate this effect (Luhmann & Eid, 2009).                                                                                                                                                                                                                  |                                                                      |                                    |                                                                                |     |
| B          | Together, these studies indicate that major life events can have strong effects on SWB and that the strength of these effects varies, depending on the life events considered.                                                                                                                                                                                                                                                                                                                         |                                                                      |                                    |                                                                                |     |
| B1         | major life events can have strong effects on SWB                                                                                                                                                                                                                                                                                                                                                                                                                                                       |                                                                      | Life Event -c→ SWB                 |                                                                                |     |
| B2         | the strength of these effects varies, depending on the life events considered                                                                                                                                                                                                                                                                                                                                                                                                                          |                                                                      | Life Event -c→ intensity of change |                                                                                |     |
| C          | SWB relates to how people feel and think about their lives.                                                                                                                                                                                                                                                                                                                                                                                                                                            | Indirectly from Diener (1984) via Luhmann et al. (2012)              | n                                  |                                                                                |     |
| D          | AWB and CWB are distinct constructs                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | Indirectly from Lucas, Diener & Suh (1996) vie Luhmann et al. (2012) |                                    |                                                                                |     |
| D1         | Affective well-being (AWB) refers to the presence of pleasant affect (e.g., feelings of happiness) and the absence of unpleasant affect (e.g., depressed mood).                                                                                                                                                                                                                                                                                                                                        |                                                                      | n                                  |                                                                                |     |
| D2         | Cognitive well-being (CWB) refers to the cognitive evaluation of life overall (i.e., global life satisfaction) as well as of specific life domains (e.g., job satisfaction or marital satisfaction).                                                                                                                                                                                                                                                                                                   |                                                                      | n                                  |                                                                                |     |
| E          | It is therefore plausible that other external life circumstances, such as life events, have differential effects on AWB and CWB and that adaptation of AWB and CWB does not occur at the same rate.                                                                                                                                                                                                                                                                                                    | Luhmann et al. (2012)                                                |                                    | Having AWB and CWB separated in the model could be a good idea                 |     |
| F          | Life events have been examined from two major perspectives (Filipp & Aymanns, 2009): a stress perspective and a developmental perspective.                                                                                                                                                                                                                                                                                                                                                             | Luhmann et al. (2012)                                                |                                    |                                                                                |     |
| F1         | From the stress perspective, life events are viewed as specific types of stressors (e.g., Park, 2010; Segerstrom & Miller, 2004). Stressful life events are all events that significantly disturb the daily routine (Turner & Wheaton, 1997). This definition explicitly includes desirable events such as marriage or vacation (Holmes & Rahe, 1967).                                                                                                                                                 |                                                                      | Life Event -c→ Stress              | The stress perspective is easier to model.                                     |     |
| F2         | By contrast, minor stressors such as daily hassles and uplifts (Kanner, Coyne, Schaefer, & Lazarus, 1981) are distinct from stressful life events. From the developmental perspective, life events are viewed as specific transitions. Transitions are defined as a “discontinuity in a person’s life space of which he is aware and which requires new behavioural responses” (Hopson & Adams, 1976, p. 24).                                                                                          |                                                                      |                                    |                                                                                |     |
| F3         | For the purpose of this metaanalysis, we developed a narrower definition of life events that integrates elements from both perspectives. According to our working definition, life events are time-discrete transitions that mark the beginning or the end of a specific status. A status is a nominal variable with at least two levels.                                                                                                                                                              | Luhmann et al. (2012)                                                | Life Event -c→ Status              |                                                                                |     |
| G          | the process perspective explicitly predicts the trajectory of SWB over time within individuals                                                                                                                                                                                                                                                                                                                                                                                                         | Luhmann et al. (2012)                                                |                                    |                                                                                |     |
| G1         | The adaptation process is initiated by an external stimulus (e.g., a major life event) that causes a physiological or psychological response (e.g., decreased SWB). Over time, the responsiveness diminishes and the level of SWB returns to its preevent level.                                                                                                                                                                                                                                       | Luhmann et al. (2012)                                                |                                    | IMPORTANT<br><br>This description can be used to model the adaptation process! |     |
| H          | Most major life events are—at least to some extent— controllable and predictable and can therefore be anticipated. This anticipation might cause a specific hedonic reaction even before the event occurred: If the spouse is terminally ill, the hedonic reaction to bereavement starts long before the spouse actually dies. These so-called anticipatory or lead effects can be observed months or even several years before the occurrence of the event (Clark, Diener, Georgellis, & Lucas, 2008) | Luhmann et al. (2012)                                                |                                    |                                                                                |     |

### 2024-03-18 manuscript.Rmd
Continue writing the text and write simulation code in R

@@
 # Seed for random number generation
-set.seed(42)
+set.seed(1256)
 knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
@@
 One of the most important questions we might attempt to answer is how to live our lives and what makes life good.
-Trying to maximize our happiness is the naive answer we easily arrive at, we want to feel good, but *happiness* is an ambiguous word.
+Trying to maximize our happiness is the naive answer we easily arrive at, we want to feel good, and it is the self-evident reality all of us would agree on: being happy feels good.
+*Happiness* is an ambiguous word.
 Although it seems like all people tend to evaluate their lives and think about *happiness* [@veenhoven11HowUniversal2010], the understanding of *happiness*, its use and connotations vary across time and place [@oishiConceptsHappinessTime2013].
@@
 ## Subjective Well-Being
-According to @dienerSubjectiveWellbeing1984, SWB is the subjective global evaluation of life and consists of three components: (1) life satisfaction, (2) positive affect, (3) negative affect.
+According to @dienerSubjectiveWellbeing1984, SWB is the subjective global evaluation of life and consists of three components: (1) life satisfaction, (2) positive affect, (3) negative affect (see the *Constructs* section for a precise definition).
 
@@
 
-The observation, that people go back to a stable level of SWB after (major) life events is not new.
+The observation, that people go back to a stable neutral level of SWB after (major) life events is not new.
 @helsonAdaptationlevelTheoryExperimental1964 formulated the more general Adaptation Level Theory, stating that how people perceive stimuli is influenced by their past experiences.
 @brickman1971adaptation applied this idea to SWB for the first time and coined the term *hedonic treadmill*, which is still commonly used to refer to the phenomenon of returning to a base SWB level after life events [e.g.
-@czaikaMigrationAspirationsAre2014, manciniSteppingHedonicTreadmill2011], and @brickmanLotteryWinnersAccident provided first evidence for it.
+@czaikaMigrationAspirationsAre2014; @manciniSteppingHedonicTreadmill2011], and @brickmanLotteryWinnersAccident provided first evidence for it.
 
@@
 
-We aim to formalize the Set-Point Theory of SWB as described by @dienerHedonicTreadmillRevising2006a, which explains the phenomenon of hedonic treadmill and tries to account for individual differences.
+We aim to formalize the *Set Point Theory of SWB* as described by @dienerHedonicTreadmillRevising2006a, which explains the phenomenon of hedonic treadmill and tries to account for individual differences.
 
@@
 So @luhmannSubjectiveWellbeingAdaptation2012 used a method that takes time into account.
-For each life event, they fitted a logarithmic model using the effects sizes reported in original studies.
+For each life event, they fitted a logarithmic model using the effect sizes reported in original studies.
 An adaptation is characterized by a change parameter, which is in the opposite direction of the initial reaction.
@@
 ### Treatment
-Studies focus on different major life events, 8 of them were included in the meta-analysis.
+Studies focus on different major life events, 8 life events were included in the meta-analysis.
 
@@
 
-# The Prototheory
+# The Theory
+
+@dienerHedonicTreadmillRevising2006a evaluated the empirical findings from the last decades and expanded the model of hedonic treadmill, creating the *Set Point Theory of SWB*.
+They made 5 adjustments:
+
+1. Set points can be non-neutral.
+2. There are inter-individual differences in set points.
+3. There are multiple set points: e.g. independent set  points for positive affect, negative affect or domain-specific well-being variables like work satisfaction, home satisfaction.
+4. Individual set points of SWB can and do change over time.
+5. There are inter-individual and intra-individual (based on the life event) differences in adaptation.
+
+The general idea is that people have (genetically inherited to an extent) set points of SWB they return to after big changes in their lives.
+In this formalization, we are only going to take adjustments (1), (2) and (5) into account, since (3) and (4) add non-negligible complexity to the model, which is out of the scope of this project.
 
@@
 
-*TODO*
+### SWB
 
-- SWB: Definiton by Diener (include subtypes in the theoretical model but not in)
-  - AWB:
-  - CWB:
+![](material/setpoint_definitions_swb.drawio.png)
+
 - SWBSetPoint: SWB people return to
@@
   - AdaptationRatePerson: Personenvariable
-  - AdaptationRateEvent: Hängt von InitialReaction ab
 - InitialReaction: Change in SWB directly after / at the time of the LifeEvent
@@
 
-# Model
-
-The central idea of
+## Model
 
@@
 
-- The constructs of *Life Event* and *Initial Reaction* can be reduced to *Initial Reaction*, since only the latter is a part of the (following) relevant psychological processes.
--
+- We do not take anticipatory effects into account.
+SWB tends to change before major life events, if people know that something is going to happen.
+Before marriage as life event people can already anticipate their wedding, which leads to an increase in SWB.
+- AWB and CWB are summarized into a single SWB construct.
+- Life Event and Initial Reaction are reduced to Initial Reaction (appraisal of the situation), since only the latter is important for the following psychological processes.
 
@@
 ```
+
+# Simulation
+```{r, include=FALSE}
+source("simulation/simulation.R")
+```
+
+## Positive Life Events
+
+```{r positive_life_event}
+plot_single_person(5, 0.2)
+```
+
+## Negative Life Events
+
+```{r negative_life_event}
+plot_single_person(5, -0.2)
+```
+
+## Testing the hypothesis
+
+```{r}
+n <- 2000
+years <- 5
+life_event <- 0.2
+
+results <- test_adaptation(n, years, life_event)
+p <- results$p.value
+```
+
+We calculated a paired sample t-test with `r n` simulated persons.
+They experienced a life event with an objective value of `r life_event`.
+We compared the SWB values just before the event and after `r years` years.
+
+$H_0$: Means of SWB values are the same before and `r years` after the event.
+
+$H_1$: Means of SWB values are different before and `r years` after the event.
+
+The t-test delivered a p-value of `r p`, which means that there are no differences in the SWB value before and 5 years after the event.
 

### 2024-03-18 material/variable_definitions.Rmd
Continue writing the text and write simulation code in R

@@
 | --------- | ----------- | ----- | ------- |
-| Subjective Well-Being | continuous | [0; 1] | 0 = absolutely not satisfied with life \newline 1 = maximal life satisfaction |
-| Subjective Well-Being Set Point | continuous | [0; 1] | 0 = absolutely not satisfied with life \newline 1 = maximal life satisfaction |
-| Appraisal of Life Event | continuous | [-1; 1] | -1 = negative \newline 0 = neutral \newline 1 = positive |
+| SWB | continuous | [0; 1] | 0 = absolutely not satisfied with life \newline 1 = maximal life satisfaction |
+| SWBSetPoint | continuous | [0; 1] | 0 = absolutely not satisfied with life \newline 1 = maximal life satisfaction |
+| InitialReaction | continuous | [-1; 1] | -1 = negative \newline 0 = neutral \newline 1 = positive |
+| AdaptationRate | continuous | [0; 1] | 0 = no adaptation \newline 1 = fastest possible adaptation |
+| AdaptationRatePerson | continous | [0; 1] | 0 = lowest possible tendency for adaptation \newline 1 = highest possible tendency for adaptation |

### 2024-03-18 simulation/simulation.R

@@ -1,0 +1,101 @@
+
+# A person is represented by a list:
+# list(sensitivity, adaptation_rate, SWB_set_point)
+
+# ====== Relationships ======
+
+initial_reaction <- function (life_event, sensitivity) {
+    life_event * sensitivity
+}
+
+delta_SWB <- function (SWB, SWB_set_point) {
+    SWB - SWB_set_point
+}
+
+adaptation_rate <- function(delta_SWB_arg, adaptation_rate_person) {
+  (-delta_SWB_arg) * adaptation_rate_person
+}
+
+SWB <- function (SWB, adaptation_rate) {
+  SWB + adaptation_rate
+}
+
+event <- function(persons, life_event) {
+  persons[, SWB := initial_reaction(life_event, sensitivity) + SWB]
+}
+
+SWB_chain <- function(person) {
+  delta_SWB_var <- delta_SWB(person$SWB, person$SWB_set_point)
+  adaptation_rate_var <- adaptation_rate(delta_SWB_var, person$adaptation_rate)
+  SWB_var <- SWB(person$SWB, adaptation_rate_var)
+  
+  person$SWB <- SWB_var
+  person
+}
+
+adaptation <- function(persons, years) {
+  for(i in seq_len(years)) {
+    # Take current row (.SD) as list and use it for the SWB_chain function
+    persons[, SWB := SWB_chain(as.list(.SD))$SWB]
+  }
+}
+
+
+# ====== Simulation ======
+
+library("data.table")
+
+test_adaptation <- function (n, years, life_event) {
+  data <- data.table(
+    sensitivity = runif(n, 0.5, 1),
+    adaptation_rate = runif(n, 0.4, 0.6),
+    SWB_set_point = runif(n, 0, 1)
+  )
+  
+  data[, SWB := SWB_set_point]
+  
+  SWB_pre <- data$SWB
+  
+  data |>
+    event(0.2) |>
+    adaptation(years)
+  
+  SWB_post <- data$SWB
+  
+  # Test
+  
+  t.test(SWB_pre, SWB_post)
+}
+
+# Plot a single person
+
+plot_single_person <- function (years, life_event) {
+  person <- list(
+    sensitivity = runif(1, 0.5, 1),
+    adaptation_rate = runif(1, 0.4, 0.6),
+    SWB_set_point = runif(1, 0, 1)
+  )
+  
+  person$SWB <- person$SWB_set_point + initial_reaction(life_event, person$sensitivity)
+  
[…]

### 2024-03-19 manuscript.Rmd
Continue working on the project

@@
 
-@luhmannSubjectiveWellbeingAdaptation2012 has done a meta-analysis on adaptation to life events focusing on SWB.
+@luhmannSubjectiveWellbeingAdaptation2012 has done a meta-analysis on adaptation to life events focusing on SWB, they included 313 studies focusing on 8 different major life events.
+The total number of participants was 65911.
 Evaluating the robustness of the studies focusing on SWB and adaptation is difficult, since it is impossible to manipulate life events experimentally and cross-sectional studies have a lack of internal validity [@luhmannSubjectiveWellbeingAdaptation2012].
@@
 
-Studies focusing on the hedonic treadmill can not be adequatly characterized by single effect sizes, since the temporal development of SWB is of interest.
+Studies focusing on the hedonic treadmill can not be adequately characterized by single effect sizes, since the temporal development of SWB is of interest.
 So @luhmannSubjectiveWellbeingAdaptation2012 used a method that takes time into account.
@@
 
+The visual representations of construct definitions and their relationships with each other according to the theory are done using VAST [@leisingVisualArgumentStructure2023].
+
 ## Constructs
+
+The central constructs of the *Set Point Theory* are SWB, set point of SWB, life event, initial reaction and adaptation rate.
+We added the constructs emotional reactivity, and personal adaptation rate.
+The VAST displays for the construct definitions are shown below.
 
@@
 
-![](material/setpoint_definitions_swb.drawio.png)
+![](material/setpoint_definitions-SWB.drawio.png)
 
-- SWBSetPoint: SWB people return to
-- AdaptationRate: How fast the adaptation happens
-  - AdaptationRatePerson: Personenvariable
-- InitialReaction: Change in SWB directly after / at the time of the LifeEvent
-- LifeEvent: Status change
+### Other relevant constructs
 
-## Model
+![](material/setpoint_definitions-Others.drawio.png)
 
-# Implementation
+We made small changes to the theoretical model, in order to make it suitable for an implementation.
 
-We made small changes to the theoretical model, in order to make it suitable for an implementation. We are making the following assumptions:
+@luhmannSubjectiveWellbeingAdaptation2012 clearly state that the adaptation after negative life events is slower. This would complicate the model and it is out of the scope of this project.
 
-- We do not take anticipatory effects into account.
+We also do not take anticipatory effects into account.
 SWB tends to change before major life events, if people know that something is going to happen.
-Before marriage as life event people can already anticipate their wedding, which leads to an increase in SWB.
-- AWB and CWB are summarized into a single SWB construct.
-- Life Event and Initial Reaction are reduced to Initial Reaction (appraisal of the situation), since only the latter is important for the following psychological processes.
+For example, before their marriage people can already anticipate their wedding, which leads to an increase in SWB.
+This is very well reflected in the empirical findings but it is not a part of the *Set Point Theory*.
 
-*VAST display of the model here*
+Empirical findings show that AWB and CWB are affected separately by life events, their adaptations are also independent.
+@dienerHedonicTreadmillRevising2006a also have this in their models, as extension (3) but they do not state how separate well-being variables are affected by life events, how different their set points are, and what they change in the adaptation process.
+
+![](material/setpoint_definitions-Theory.drawio.png)
 
@@
 
-# Simulation
+# Simulation and Evaluation of the Model
 ```{r, include=FALSE}
@@
 :::
+
+# (APPENDIX) Appendix {-}
+
+# Statements
+
+```{r child="./material/statements.Rmd"}
+```

### 2024-03-19 material/statements.Rmd
Continue working on the project

@@
-| Identifier | Citation                                                                                                                                                                                                                                                                                                                                                                                                                                                                                               | Reference                                                            | Relationship?
-| Comments?
-|     |
-| ---------- | ------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------ | -------------------------------------------------------------------- | ---------------------------------- | ------------------------------------------------------------------------------ | --- |
+- ”MA” (meta-analysis): Luhmann et al. (2012)
+- “TP” (theory paper): Diener et al. (2006)
+- “OI” (original idea): Ideas / assumptions I added myself.
+- “OI(MA/TP/MA+TP)”: Original ideas I extracted from things mentioned  in the meta-analysis or theory paper. It is also possible to combine insights from both papers.
+
+| ID  | Citation                                                                                                                                                                                                                                                                                                                                                                                                                                                                                               | Reference                                    |
+| --- | ------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------ | -------------------------------------------- |
 | A   | A notable finding was that the initial reaction and the rate of adaptation varied considerably between different life events.
-| Luhmann et al.
-(2012)                                                | Appraisal  - c → Adaptation Rate   | Valence of the life event changes the adaptation rate.
-|     |
+| MA                                           |
 | A1  | For instance, the initial reaction to marriage was positive, but subsequent adaptation was fast and completed after 2 years on average (Lucas et al., 2003).
-|                                                                      |                                    | positive valence: fast adaptation                                              |     |
+| MA                                           |
 | A2  | In contrast, the rate of adaptation was much slower for negative events such as the onset of disability (Lucas, 2007b), widowhood (Lucas et al., 2003; Specht, Egloff, & Schmukle, 2011), and divorce (Lucas, 2005).
-|                                                                      |                                    | negative valence: slow adaptation                                              |     |
+| MA                                           |
 | A3  | The findings were perhaps most striking for unemployment: After the beginning of unemployment, mean levels of SWB were significantly below baseline, even if people became reemployed (Lucas et al., 2004).
 Repeated unemployment spells aggravate this effect (Luhmann & Eid, 2009).
-|                                                                      |                                    |                                                                                |     |
+| MA                                           |
 | B   | Together, these studies indicate that major life events can have strong effects on SWB and that the strength of these effects varies, depending on the life events considered.
-|                                                                      |                                    |                                                                                |     |
-| B1         | major life events can have strong effects on SWB                                                                                                                                                                                                                                                                                                                                                                                                                                                       |                                                                      | Life Event -c→ SWB                 |                                                                                |     |
-| B2         | the strength of these effects varies, depending on the life events considered                                                                                                                                                                                                                                                                                                                                                                                                                          |                                                                      | Life Event -c→ intensity of change |                                                                                |     |
+| MA                                           |
+| B1  | major life events can have strong effects on SWB                                                                                                                                                                                                                                                                                                                                                                                                                                                       | MA                                           |
+| B2  | the strength of these effects varies, depending on the life events considered                                                                                                                                                                                                                                                                                                                                                                                                                          | MA                                           |
 | C   | SWB relates to how people feel and think about their lives.
-| Indirectly from Diener (1984) via Luhmann et al.
-(2012)              | n                                  |                                                                                |     |
-| D          | AWB and CWB are distinct constructs                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | Indirectly from Lucas, Diener & Suh (1996) vie Luhmann et al.
-(2012) |                                    |                                                                                |     |
+| Indirectly Diener (1984) via MA              |
+| D   | AWB and CWB are distinct constructs                                                                                                                                                                                                                                                                                                                                                                                                                                                                    | Indirectly Lucas, Diener & Suh (1996) via MA |
 | D1  | Affective well-being (AWB) refers to the presence of pleasant affect (e.g., feelings of happiness) and the absence of unpleasant affect (e.g., depressed mood).
-|                                                                      | n                                  |                                                                                |     |
+| MA                                           |
 | D2  | Cognitive well-being (CWB) refers to the cognitive evaluation of life overall (i.e., global life satisfaction) as well as of specific life domains (e.g., job satisfaction or marital satisfaction).
-|                                                                      | n                                  |                                                                                |     |
+| MA                                           |
 | E   | It is therefore plausible that other external life circumstances, such as life events, have differential effects on AWB and CWB and that adaptation of AWB and CWB does not occur at the same rate.
-| Luhmann et al.
-(2012)                                                |                                    | Having AWB and CWB separated in the model could be a good idea                 |     |
+| MA                                           |
 | F   | Life events have been examined from two major perspectives (Filipp & Aymanns, 2009): a stress perspective and a developmental perspective.
-| Luhmann et al.
-(2012)                                                |                                    |                                                                                |     |
+| MA                                           |
 | F1  | From the stress perspective, life events are viewed as specific types of stressors (e.g., Park, 2010; Segerstrom & Miller, 2004).
@@
 This definition explicitly includes desirable events such as marriage or vacation (Holmes & Rahe, 1967).
-|                                                                      | Life Event -c→ Stress              | The stress perspective is easier to model.
-|     |
+| MA                                           |
 | F2  | By contrast, minor stressors such as daily hassles and uplifts (Kanner, Coyne, Schaefer, & Lazarus, 1981) are distinct from stressful life events.
@@
 24).
-|                                                                      |                                    |                                                                                |     |
+| MA                                           |
 | F3  | For the purpose of this metaanalysis, we developed a narrower definition of life events that integrates elements from both perspectives.
 According to our working definition, life events are time-discrete transitions that mark the beginning or the end of a specific status.
-A status is a nominal variable with at least two levels.
-| Luhmann et al.
-(2012)                                                | Life Event -c→ Status              |                                                                                |     |
-| G          | the process perspective explicitly predicts the trajectory of SWB over time within individuals                                                                                                                                                                                                                                                                                                                                                                                                         | Luhmann et al.
-(2012)                                                |                                    |                                                                                |     |
+| MA                                           |
+| F4  | A status is a nominal variable with at least two levels.
+|                                              |
+| G   | the process perspective explicitly predicts the trajectory of SWB over time within individuals                                                                                                                                                                                                                                                                                                                                                                                                         | MA                                           |
 | G1  | The adaptation process is initiated by an external stimulus (e.g., a major life event) that causes a physiological or psychological response (e.g., decreased SWB).
 Over time, the responsiveness diminishes and the level of SWB returns to its preevent level.
-| Luhmann et al.
-(2012)                                                |                                    | IMPORTANT<br><br>This description can be used to model the adaptation process!
-|     |
+| MA                                           |
 | H   | Most major life events are—at least to some extent— controllable and predictable and can therefore be anticipated.
 This anticipation might cause a specific hedonic reaction even before the event occurred: If the spouse is terminally ill, the hedonic reaction to bereavement starts long before the spouse actually dies.
-These so-called anticipatory or lead effects can be observed months or even several years before the occurrence of the event (Clark, Diener, Georgellis, & Lucas, 2008) | Luhmann et al.
-(2012)                                                |                                    |                                                                                |     |
+These so-called anticipatory or lead effects can be observed months or even several years before the occurrence of the event (Clark, Diener, Georgellis, & Lucas, 2008) | MA                                           |
+| I   | The original treadmill theory suggested that people return to a neutral set point after an emotionally significant event.
+However, decades of research show that this part of the hedonic treadmill theory is wrong.
+Instead, most people are happy most of the time (Diener & Diener, 1996).
+| TP                                           |
+| J   | The empirical research that has been conducted since the publication of Brickman and Campbell (1971) reveals that if people do have set points, they vary considerably across individuals.
+| TP                                           |
+| K   | But we have found individual differences in the rate and extent of adaptation that occurs even to the same event.
+In our longitudinal studies, the size and even the direction of the change in life satisfaction varied considerably across individuals.
+| TP                                           |
+| L   | (not a direct citation) Set point of SWB = the level of SWB people return to after life events.
+| TP                                           |
+| M   | The difference between the current SWB and the set point affects the change rate.
+If the difference is larger, expected change is also larger.
+| OI(MA+TP)                                    |
+| N   | (not a direct citation) People react differently to life events                                                                                                                                                                                                                                                                                                                                                                                                                                        | MA                                           |

### 2024-03-19 material/variable_definitions.Rmd
Continue working on the project

--- 2024-03-16 material/variable_definitions.Rmd
@@
 | --------- | ----------- | ----- | ------- |
-| Subjective Well-Being | continuous | [0; 1] | 0 = absolutely not satisfied with life \newline 1 = maximal life satisfaction |
-| Subjective Well-Being Set Point | continuous | [0; 1] | 0 = absolutely not satisfied with life \newline 1 = maximal life satisfaction |
-| Appraisal of Life Event | continuous | [-1; 1] | -1 = negative \newline 0 = neutral \newline 1 = positive |
+| SWB | continuous | [0; 1] | 0 = absolutely not satisfied with life \newline 1 = maximal life satisfaction |
+| SWBSetPoint | continuous | [0; 1] | 0 = absolutely not satisfied with life \newline 1 = maximal life satisfaction |
+| LifeEvent | continous | [-1; 1] | -1 = objectively very negative \newline 0 = objectively neutral \newline 1 = objectively very positive |
+| EmotionalReactivity | continuous | [0; 1] | 0 = emotional state is not influenced by external events \newline 1 = external events influence emotional states very strongly |
+| InitialReaction | continuous | [-1; 1] | -1 = negative \newline 0 = neutral \newline 1 = positive |
+| AdaptationRate | continuous | [0; 1] | 0 = no adaptation \newline 1 = fastest possible adaptation |
+| AdaptationRatePerson | continous | [0; 1] | 0 = lowest possible tendency for adaptation \newline 1 = highest possible tendency for adaptation |

### 2024-03-19 simulation/simulation.R

@@ -1,13 +1,66 @@
 
 # A person is represented by a list:
-# list(sensitivity, adaptation_rate, SWB_set_point)
+# list(emotional_reactivity, adaptation_rate, SWB_set_point)
+
+# ====== Helpers ======
+
+library(dplyr)
+
+#' A generic logistic function
+#' 
+#' @param x 
+#' @param d
+#' @param a
+#' @param min
+#' @param max
+logistic <- function (x, d = 0, a = 1, min = 0, max = 1) {
+  min + (max - min)/(1 + exp(a * (d - x)))
+}
+
+#' Logistic function that delivers values between 0 - 1 for the interval 0 - 1
+#'
+#' @param x
+logistic_0_1 <- function(x) {
+  logistic(x, d = 0.5, a = 10)
+}
+
+#' Function representing a linear relationship.
+#' Overflows are prevented.
+#' 
+#' @param x
+#' @param min_y Minimum y value
+#' @param max_y Maximum y value
+#' @param min_x Minimum y value
+#' @param max_x Maximum y value
+#' 
+limited_linear <- function(
+    x,
+    min_y = 0, max_y = 1,
+    min_x = 0, max_x = 1
+  ) {
+  slope = (max_y - min_y) / (max_x - min_x)
+  
+  result <- case_when(
+    x > max_x ~ max_y,
+    x < min_x ~ min_y,
+    TRUE ~ slope * x
+  )
+}
 
 # ====== Relationships ======
 
-initial_reaction <- function (life_event, sensitivity) {
-    life_event * sensitivity
+#' Calculate initial emotional reaction of the person
+#' 
+#' @param life_event Objective valence of the life event
+#' @param  emotional_reactivity Tendency of the person to be influenced emotionally by external events
+initial_reaction <- function (life_event, emotional_reactivity) {
+    life_event * emotional_reactivity
 }
 
+#' Calculate the difference between the SWB and set point
+#' 
+#' @param SWB Current SWB
+#' @param SWB_set_point Baseline level of SWB for a person
 delta_SWB <- function (SWB, SWB_set_point) {
     SWB - SWB_set_point
 }
@@ -17,11 +70,11 @@ adaptation_rate <- function(delta_SWB_arg, adaptation_rate_person) {
 }
 
 SWB <- function (SWB, adaptation_rate) {
-  SWB + adaptation_rate
+  limited_linear(SWB + adaptation_rate)
 }
 
[…]

### 2024-03-19 manuscript.Rmd
Add documentation to code

@@
 ```{r positive_life_event}
-plot_single_person(5, 0.2)
+person <- list(
+  emotional_reactivity = 0.5,
+  adaptation_rate = 0.5,
+  SWB_set_point = 0.5
+)
+
+plot_single_person(person, 5, 0.2)
 ```
@@
 ```{r negative_life_event}
-plot_single_person(5, -0.2)
+person <- list(
+  emotional_reactivity = 0.5,
+  adaptation_rate = 0.5,
+  SWB_set_point = 0.5
+)
+
+plot_single_person(person, 5, -0.2)
 ```

### 2024-03-19 simulation/simulation.R

@@ -1,10 +1,10 @@
-
-# A person is represented by a list:
-# list(emotional_reactivity, adaptation_rate, SWB_set_point)
-
-# ====== Helpers ======
+# ====== Packages ======
 
 library(dplyr)
+library(data.table)
+library(ggplot2)
+
+# ====== Helpers ======
 
 #' A generic logistic function
 #' 
@@ -65,18 +65,26 @@ delta_SWB <- function (SWB, SWB_set_point) {
     SWB - SWB_set_point
 }
 
+#' Calculate the adaptation rate for a specific person at a specific delta SWB
+#' 
+#' @param delta_SWB_arg Current delta SWB
+#' @param adaptation_rate_person Personal adaptation rate
 adaptation_rate <- function(delta_SWB_arg, adaptation_rate_person) {
   (-delta_SWB_arg) * adaptation_rate_person
 }
 
+#' Calculate the SWB given current SWB and adaptation rate
+#' 
+#' @param SWB Current SWB
+#' @param adaptation_rate Adaptation rate at the moment
 SWB <- function (SWB, adaptation_rate) {
   limited_linear(SWB + adaptation_rate)
 }
 
-event <- function(persons, life_event) {
-  persons[, SWB := initial_reaction(life_event, emotional_reactivity) + SWB]
-}
 
+#' Calculate the SWB for a given person for the next iteration
+#' 
+#' @param person A list representing a person
 SWB_chain <- function(person) {
   delta_SWB_var <- delta_SWB(person$SWB, person$SWB_set_point)
   adaptation_rate_var <- adaptation_rate(delta_SWB_var, person$adaptation_rate)
@@ -86,6 +94,18 @@ SWB_chain <- function(person) {
   person
 }
 
+#' Calculate the SWB for multiple persons after a life event
+#' 
+#' @param persons A data.table containing the persons
+#' @param life_event Objective value of a life event
+event <- function(persons, life_event) {
+  persons[, SWB := initial_reaction(life_event, emotional_reactivity) + SWB]
+}
+
+#' Simulate an adaptation
+#' 
+#' @param persons A data.table containing the persons
+#' @param years Years to wait
 adaptation <- function(persons, years) {
   for(i in seq_len(years)) {
     # Take current row (.SD) as list and use it for the SWB_chain function
@@ -96,9 +116,16 @@ adaptation <- function(persons, years) {
 
 # ====== Simulation ======
 
-library("data.table")
[…]

### 2024-03-19 manuscript.Rmd
Finalize everything

@@
 abstract: |
-  There is research that shows...
-
-  Here, we model the set point theory of subjective well being
+  We attempted to model the Set Point Theory of Subjective Well-Being by Diener et al.
+(2006).
+We used VAST by Leising et al.
+(2023) to visualize definitions and relationships of constructs.
+We finally implemented the resulting mathematical model using the programming language R.
 
@@
 
-source("VAST/VAST.R")
+source("simulation/simulation.R")
 ```
@@
 ### Setting
-aedasd
+No clear statements are possible based on the meta-analysis.
 
@@
 
+\newpage
+
+## Relationships between Constructs
+
+### InitialReaction
+
+The calculation of the initial reaction is based on the objective value of the life event and the emotional reactivity of the person.
+
+$$\text{InitialReaction} = \text{LifeEvent} \times \text{EmotionalReactivity}$$
+```{r}
+# Define emotional reactivity sequence
+emotional_reactivity_seq <- seq(0, 1, 0.2)
+
+# Create a grid for plotting
+par(mfrow = c(2, 3))  # 2 rows, 3 columns
+
+# Plot the function for each emotional reactivity value
+for (i in 1:length(emotional_reactivity_seq)) {
+    emotional_reactivity <- emotional_reactivity_seq[i]
+    life_events <- seq(-1, 1, length.out = 100)  # Generating life events in the range [-1, 1]
+    reactions <- initial_reaction(life_events, emotional_reactivity)  # Calculating reactions
+    plot(life_events, reactions, type = 'l', main = paste("EmotionalReactivity:", emotional_reactivity),
+         xlab = "LifeEvent", ylab = "InitialReaction")
+}
+```
+
+### DeltaSWB
+
+DeltaSWB is simply the difference between the personal SWB set point and the current SWB.
+
+$$\text{DeltaSWB} = \text{SWB} - \text{SWBSetPoint}$$
+
+```{r}
+# Define SWB_set_point sequence
+SWB_set_points <- seq(0, 1, by = 0.2)
+
+# Create a grid for plotting
+par(mfrow = c(2, 3))  # 2 rows, 3 columns
+
+# Plot the function for each SWB_set_point
+for (i in 1:length(SWB_set_points)) {
+    SWB_set_point <- SWB_set_points[i]
+    SWB_values <- seq(0, 1, length.out = 100)  # Generating SWB values
+    deltas <- delta_SWB(SWB_values, SWB_set_point)  # Calculating deltas
+    plot(SWB_values, deltas, type = 'l', main = paste("SWBSetPoint:", SWB_set_point),
+         xlab = "SWB", ylab = "DeltaSWB")
+}
+```
+
+### Adaptation Rate
+
+AdaptationRate is based on the interaction of DeltaSWB and AdaptationRatePerson.
+We want to have a logarithmic result at the end [@luhmannSubjectiveWellbeingAdaptation2012], using the current DeltaSWB for each iteration gives us exactly that.
+The adaptation should happen in the opposite direction of the InitialReaction, so we start by negating DeltaSWB (-DeltaSWB).
+According to @dienerHedonicTreadmillRevising2006a, there are individual differences in the adaptation rate, so we multiply the DeltaSWB based change with AdaptationRatePerson.
+
+$$\text{AdaptationRate} = -\text{DeltaSWB} \times \text{AdaptationRatePerson}$$
+```{r}
+# Define adaptation_rate_person sequence
+adaptation_rates <- seq(0, 1, by = 0.2)
+
+# Create a grid for plotting
+par(mfrow = c(2, 3))  # 2 rows, 3 columns
+
+# Plot the function for each adaptation_rate_person value
+for (i in 1:length(adaptation_rates)) {
+    adaptation_rate_person <- adaptation_rates[i]
+    delta_SWB_values <- seq(-1, 1, length.out = 100)  # Generating delta_SWB values
+    rates <- adaptation_rate(delta_SWB_values, adaptation_rate_person)  # Calculating adaptation rates
+    plot(delta_SWB_values, rates, type = 'l', main = paste("AdaptationRatePerson:", adaptation_rate_person),
+         xlab = "DeltaSWB", ylab = "AdaptationRate")
+}
+```
+
+### SWB
+
+For the calculation of SWB, we just add AdaptationRate to the previous SWB value. We also limit the resulting values to the valid range for SWB, which is [0, 1].
+
+$$
+\text{SWB}_{\text{next}}(x) = \begin{cases} 
+      0 & \text{if } (\text{SWB}_{\text{old}} + \text{AdaptationRate}) < 0 \\
+      \text{SWB}_{\text{old}} + \text{AdaptationRate} & \text{if } 0 < (\text{SWB}_{\text{old}} + \text{AdaptationRate}) < 1 \\
+      1 & \text{if } (\text{SWB}_{\text{old}} + \text{AdaptationRate}) > 1
+\end{cases}
+$$
+```{r}
+# Generate SWB sequence
+SWB_values <- seq(0, 1, by = 0.2)
+
+# Define adaptation_rate sequence
+adaptation_rates <- seq(-1, 1, by = 0.2)
+
+# Create a grid for plotting
+par(mfrow = c(2, 3))  # 2 rows, 3 columns
+
+# Plot the function for each adaptation_rate
+for (i in 1:length(adaptation_rates)) {
+    adaptation_rate <- adaptation_rates[i]
+    SWB_result <- SWB(SWB_values, adaptation_rate)
+    plot(SWB_values, SWB_result, type = 'l', main = paste("Adaptation Rate:", adaptation_rate),
+         xlab = "SWB", ylab = "SWB with Adaptation")
+}
+
+```
+
 # Simulation and Evaluation of the Model
-```{r, include=FALSE}
+
+We implemented the model using the programming language R [@R-base].
+
+```{r include=FALSE}
 source("simulation/simulation.R")
@@
 ## Positive Life Events
+
+We simulated the adaptation process of a person with an emotional reactivity of 0.5, an adaptation rate of 0.5 and a set point of 0.5.
+The objective value of the life event was set to 0.2.
+Following plot is a visualization of the adaptation process during the following 5 years after the event.
 
@@
 
+We simulated the adaptation process of a person with an emotional reactivity of 0.5, an adaptation rate of 0.5 and a set point of 0.5.
+The objective value of the life event was set to -0.2.
+Following plot is a visualization of the adaptation process during the following 5 years after the event.
+
 ```{r negative_life_event}
@@
 
-## Testing the hypothesis
+## Do People Return to the Same SWB after 5 Years?
 
@@
 
+I think TCM and other approaches we had in general worked reasonably well for the Set Point Theory.
+But in order to be able to model all kinds of theories, we need a lot more tools than we had a chance to learn about during Empra.
+The Set Point Theory and Hedonic Treadmill were pushing the limits of the tools and frameworks we learned about.
+Because of the impossibility of experimental studies in this field of research, using our methodology for evaluating the robustness of the phenomenon was really hard.
+
+The hardest step for me was getting to understand the concepts.
+I had to read a lot in order to understand what it was all about.
+I would need a lot more time to really understand everything.
+The meta-analysis I chose also used more complex statistical methods than I am used to, I decided to use it without understanding everything.
+My theory paper was really clear about the revisions to the previous ideas and the meta-analysis had definitions in the constructs in it.
+But none of the papers I read describe the exact processes behind adaptation, so I had to add lots of new constructs like EmotionalReactivity or DeltaSWB to make it work as described in the papers I read.
+
+One point I had problems with was defining constructs.
+"Do we have to define all words that are present in definitions" was an important questions.
+If you try to do it, it ends in an infinite regress, where we need more and more definitions.
+I decided to leave out definitions of "Positive Affect", "Negative Affect" and "Life Satisfaction".
+Defining "Life Satisfaction" would require a completely different literature review, I would happily go down the rabbit hole looking to understand life satisfaction, but it would be too much for this course.
+
 \newpage
